Flooding of the Okavango Delta influences Connectivity for Dispersing African Wild Dogs

### 2023-01-04 Chapter_2/04_Manuscript/Manuscript.tex
Various updates

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
\usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}

@@

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Dispersal Through Seasonal Landscapes}

% \title{Step by Step: The Utility of Dispersal Simulations to Assess Landscape
% Connectivity}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,\S} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \and
  Gabriele Cozzi\textsuperscript{1,2} \and
  Dominik M. Behr\textsuperscript{1,2}
}

@@

\textbf{Running Title:} Release the Dogs! Simulating Wild Dog Dispersal to
Assess Landscape Connectivity

@@

\textbf{Keywords:} dispersal, simulation, movement, integrated step selection,
Kavango-Zambezi Transfrontier Conservation Area, landscape connectivity, Lycaon
pictus

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
This is the abstract.
\end{abstract}

@@

% Create linenumbers
\linenumbers

\section{Introduction}

% Taken from Zeller.2020
Connectivity for wildlife is typically modeled in static environments and
estimated for a single snapshot in time (e.g., [6,7]). However, wildlife
responses to landscapes are inherently dynamic, as are the landscapes themselves
(e.g., [8])

Modeling connectivity in a static environment may mask the importance of connectivity for
maintaining wildlife population dynamics and may result in identifying corridors that are only
periodically functional [15,16,18]

@@

@@

\section{Data Availability}
GPS movement data of dispersing coalitions will be made available on dryad at
the time of publication. Access to all R-scripts for our analyses is provided
through Github.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we are indebted
to Johannes Signer for assisting with the simulation algorithm. This study was
funded by Basler Stiftung für Biologische Forschung, Claraz Foundation, Idea
Wild, Jacot Foundation, National Geographic Society, Parrotia Stiftung, Stiftung
Temperatio, Wilderness Wildlife Trust Foundation, Forschungkredit der
Universität Zürich, and a Swiss National Science Foundation Grant
(31003A\_182286) to A. Ozgul.

>>>>

\documentclass[abstract=on,10pt,a4paper,bibliography=totocnumbered]{article}
\usepackage[paper=a4paper,left=35mm,right=35mm,top=25mm,bottom=30mm]{geometry}
\usepackage[doublespacing]{setspace}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[round]{natbib}
\usepackage{amsmath}
\usepackage{colortbl}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{enumitem}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage[hidelinks]{hyperref}
\usepackage[nameinlink]{cleveref}
% \usepackage{lineno}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage[flushleft]{threeparttable}
\usepackage[nomarkers, nolists]{endfloat}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{scalerel}
\usepackage{tikz}
\usetikzlibrary{svg.path}

@@

% To be able to put an ORCID
\definecolor{orcidlogocol}{HTML}{A6CE39}
\tikzset{
  orcidlogo/.pic={
    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
  }
}
\newcommand\orcid[1]{\href{https://orcid.org/#1}{\mbox{\scalerel*{
\begin{tikzpicture}[yscale=-1,transform shape]
  \pic{orcidlogo};
\end{tikzpicture}
}{|}}}}

%------------------------------------------------------------------------------
%	Titlepage: Header
%------------------------------------------------------------------------------
\title{Flooding of the Okavango Delta influences Connectivity for Dispersing
African Wild Dogs}

% List of Authors
\author{
  David D. Hofmann\textsuperscript{1,2,\S} \orcid{0000-0003-3477-4365} \and
  Gabriele Cozzi\textsuperscript{1,2} \orcid{0000-0002-1744-1940} \and
  John W. McNutt\textsuperscript{2} \and
  Arpat Ozgul\textsuperscript{1} \orcid{0000-0001-7477-2642} \and
  Dominik M. Behr\textsuperscript{1,2} \orcid{0000-0001-7378-8538}
}

@@

\textbf{Running Title:} None

@@

\textbf{Keywords:} Seasonal Flooding of the Okavango Delta and its Consequences
for African Wild Dog Connectivity

@@

%------------------------------------------------------------------------------
%	Abstract
%------------------------------------------------------------------------------
\newpage
\begin{abstract}
Many ecosystems experience substantial changes in environmental conditions due
to seasonality. While such seasonal changes may drastically alter connectivity
for endangered species, most studies represent the environment by a static set
of spatial layers, thus ignoring seasonal variation. Here, we address this
shortcoming and employ individual-based simulations to investigate how seasonal
flooding of the Okavango delta influences connectivity for dispersing African
wild dogs (\textit{Lycaon pictus}). Our results show that the Okavango delta
poses a substantial dispersal barrier when the flood is at maximum extent, yet
that viable dispersal corridors exist when the flood is at a minimum level.
Despite a better understanding of the conservation needs for African wild dogs,
our study also provides evidence that incorporating seasonality in studies of
connectivity is imperative to more accurately predict the dispersal ability of
endangered species.
\end{abstract}

@@

% Create linenumbers
% \linenumbers

\section{Introduction}
Hello, I'm an introduction.

However, \cite{Hofmann.2021} and \cite{Hofmann.2022} used a single floodmap...

@@

\subsection{Study Area}
As depicted in \Cref{StudyArea}, the study area encomassed ...

\subsection{Spatial Habitat Layers}
We attempted to represent each spatial covariate at the highest temporal
resolution available, which resulted in various degrees of temporal accuracy. In
order to generate spatial layers that depict the flood extent at different
points in time, we ... \Cref{FloodExtent}.

\subsection{Dispersal Simulation}
We used a previously parametrized movement model to simulate dispersal of
African wild dogs. To simulate dispersing wild dogs, we employed the simulation
framework presented in \cite{Hofmann.2022}. In this framework, a movement model
that renders habitat and movement preferences of dispersing individuals is
parametrized using step-selection functions. Once parametrized, the model can be
employed to simulate virtual dispersers moving across the landscape.

We released virtual dispersers at random locations within the four distinct
source areas depicted in \Cref{StudyArea}. From each source area we simulated xx
individuals moving across the landscape at minimum flood and another xx
individuals at maximum flood. For comparison, we also ran the simulations for a
medium flood extent, yet the results for this will be presented in the appendix.

\subsection{Connectivity}
To gain insights into landscape connectivity from the simulated dispersal
events, we prepared three complementary connectivity maps (see again
\citep{Hofmann.2022}); a heatmap, indicating the absolute frequency at which
each pixel in the study area was traversed by a virtual disperser, a betweenness
map, highlighting pixels of particular importance to facilitate movement from
one area to another, and a map of interpatch connectivity, depicting the
presence and intensity of functional links between source areas. To compute the
heatmap, we counted how many times each pixel in the study area was traversed by
a virtual disperser. To calculate betweenness, we overlayed the study area with
a regular grid and used the simulated trajectories to determine movements from
one grid-cell to another. Specifically, we counted how often transitions from
one grid-cell to any other occured. The centerpoint of each grid-cell then
served as network node, and the number of transitions from one cell to another
as weights between the respective nodes. Based on the so generated network, we
then computed betweenness scores for all network nodes (i.e. for all grid-cells
across the study area). Lastly, to compute interpatch connectivity, we
determined the frequency at which virtual dispersers originating from one source
area reached any of the other source areas. We also calculated the average
dispersal duration needed until individuals reached the respective area.

@@

\section{Data Availability}
GPS movement data of dispersing wild dogs is available on dryad
\citep{Hofmann.2021b}. Access to R-scripts that exemplify the application of the
proposed approach using simulated data are provided through Github
(\url{https://github.com/DavidDHofmann/DispersalSimulation}). In addition, all
codes required to reproduce the African wild dog case study will be made
available through an online repository at the time of publication.

\section{Acknowledgements}
We thank the Ministry of Environment and Tourism of Botswana for granting
permission to conduct this research. We thank C. Botes, I. Clavadetscher, and G.
Camenisch for assisting with wild dog immobilizations. We also thank B. Abrahms
for sharing her data of three dispersing wild dogs. Furthermore, we would like
to thank Johannes Signer for assisting with the simulation algorithm. This study
was funded by Albert-Heim Foundation, Basler Stiftung für Biologische Forschung,
Claraz Foundation, Idea Wild, Jacot Foundation, National Geographic Society,
Parrotia Stiftung, Stiftung Temperatio, Wilderness Wildlife Trust Foundation,
Forschungkredit der Universität Zürich, and a Swiss National Science Foundation
Grant (31003A\_182286) to A. Ozgul.

### 2023-01-04 Chapter_8/04_Manuscript/Manuscript.tex
Various updates

Talk about connectivity models
Talk about functional connectivity
Talk about habitat fragmentation and destruction

Osipova

@@

>>>>

Talk about connectivity models
Talk about functional connectivity
Talk about habitat fragmentation and destruction
Say that wild dogs are keystone predators and an umberella species for
conservation

@@

Earlier studies demonstrating the importance of rendering seasonal change
include \cite{Chetkiewicz.2009}, who use resource selection functions to demonstrate
that connectivity differs for Ursus actor and Puma concolor and \cite{Mui.2017},
who showed that connectivity for turtles changes depending on the season, and
\cite{Kaszta.2021}, who show monthly differences in elephant's habitat selection
using path-selection functions.
